---
title: "Ireland AirBnB Analysis"
output: html_notebook
author: Nicole Dallar
---

Ireland! The Emerald Isle... Éire... The Island of Saints... Whatever you choose to call it, this beautiful land is beautiful and worthy of a visit.

I spent a semester abroad at University College Dublin during my junior year of undergrad, and I have always wanted to take another visit there now that I am employed and able to afford more than Supermac's! (But I refuse to listen to any trash talking about Milano's pizza.)

In the interest of my Irish vacation fantasy, I chose to analyze the most recent Airbnb data from their listings in Ireland in order to inform my decision of where to stay.

I want to travel on the cheaper side (hey, I'm being responsible) but definitely not as cheap as I did while an undergrad. I'd like to stay in Dublin, but I am open to staying in another county if I can get much better accomodations for the same price.

I'd also like to know which regions of Ireland have the highest ratings among Airbnb tourists. If there are any significantly better rated areas, it might be worth a visit to that area.

------

I'll start by loading in some R packages that will be used throughout the project.

```{r}
packages = c("dplyr", "ggplot2", "tidyr", "stringr")
tmp = sapply(packages, function(p){
  if (!require(p, character.only = T)){
    install.packages(p)
    library(p, character.only = T)
  }
})
```

Then I will set my working directory to my GitHub repo and read in the raw data from Airbnb.

```{r}
baseDir = normalizePath("/Users/Nicole/Documents/GitHub/IrishAirBnb")
irishRaw = read.csv(file = paste0(baseDir, "/longlistings.csv"), strip.white = T, stringsAsFactors = F, na.strings=c("","NA"))
```

Throughout this project, I found I needed to write a few custom functions. To save space and save the reader time, I have these functions written in a separate R file called "Irish_functions.R" and will source these functions into my current workspace so I can use them on the data.

```{r}
source(paste0(baseDir, "/Irish_functions.R"))
```

First I will inspect the Airbnb data (called "irishRaw") to make sure everything looks correct.

```{r}
paste0("The class of the raw data is ", class(irishRaw))
paste0("The raw data has ", dim(irishRaw)[1], " rows and ", dim(irishRaw)[2], " columns.")
```

IrishRaw is definitely a dataframe, and there are certainly many columns that might help me plan my fake vacation.

I'm going to remove some columns that I don't care about, such as URLs to images used on the listing's page or what country the listing is in (since they are all from Ireland). I also removed data that was missing values more than 30% of the listings.

```{r}
irishRaw %>% summarize_all(funs(sum(is.na(.)) / length(.)))
```

```{r}
irishSubset = irishRaw %>%
  select(-listing_url, -scrape_id, -thumbnail_url, -medium_url, -picture_url, -xl_picture_url, -host_url, -host_thumbnail_url,
         -host_picture_url, -latitude, -license, -requires_license, -weekly_price, -monthly_price, -square_feet, -jurisdiction_names,
         -longitude, -is_location_exact, -calendar_last_scraped, -country, -country_code, -availability_30, -availability_60,
         -availability_90, -host_total_listings_count)
```

One of the things that is most time-consuming about data analysis is cleaning data. Seriously, some people report spending 60% of their project time on data cleansing alone.

So, here is the first of the custom functions I mentioned earlier. I wrote a function called formatMoney that will fix the columns that are meant to be a numeric value but are irritatingly formatted with dollar symbols and commas, which stop R from knowing that it's meant to be a number. 

Here is what it looks like before:

```{r}
head(irishSubset$price)
```


```{r}
irishFixed = irishSubset %>%
  mutate_at(c("price", "security_deposit", "cleaning_fee", "extra_people"), formatMoney)

head(irishFixed$price)
```
Much better!!

Of course, there is still the issue of obscene outliers that needs to be resolved. If we stick with the example of the "price" column, certain entries seem incorrect.

```{r}
paste0("The cheapest listing costs $", min(irishFixed$price), " per night and the most expensive costs $", max(irishFixed$price), " per night.")
```

So, it seems doubtful that there are completely free listings (as much as my wallet would love that) and listings that cost more than 2,400 chicken snack boxes from Supermac's.

I'm going to use the interquartile range method to remove outliers, where anything 1.5 times the IQR past the first and third quartiles is considered an outlier. I wrote another custom function for this where a numeric column is passed to the function, which returns information about the outlier-defining limits. I chose to have it return information rather than just filter the data within the function because some human discretion can be helpful based on the column type.

```{r}
iqrRemoveOutliers(irishFixing$price)
iqrRemoveOutliers(irishFixing$cleaning_fee)
iqrRemoveOutliers(irishFixing$security_deposit)
```

Now, I know that I actually want to filter these columns so that the minimum value is greater than $0, as there is no such thing as a completely free listing on Airbnb. 

Now we are going to filter the dataframe so that the outliers are removed.

```{r}
irish = irishFixing %>%
  filter(0 < price & price < 225,
         cleaning_fee < 110,
         security_deposit < 375)
```











```{r}
by_room_type = irish %>%
  group_by(room_type) %>%
  summarize(avg_price = mean(price))
```

```{r}
ggplot(irish, aes(x = room_type, y = price)) +
  geom_boxplot()
```













